---
title: "STAT 432  Final Project"
subtitle: "Detecting Volcanoes on Venus via Classification (Where are the Volcanoes?!!)"
author: 
- 'Team : Mamamia!'
- "Di Ye (diye2), Hao Wang (haow2),Hannah Pu (chpu2)"
date: "11/6/2018"
output: 
  pdf_document:
    toc: yes
---

## Dataset Description:
The data was downloaded from Kaggle, which is originally from NASA's Magellan spacecraft database. 9734 images were captured by the spacecraft and converted to pixel (110x110, from 0 to 255), where every image is one row of 12100 columns (all the 110 rows of 110 columns). Images can contain more than one volcanoes or maybe none.
The 9000+ images are separated to two datasets (train and test):     
**Train_image** : 7000 images as train data with 12100 variables;  
**Test_image** : 2734 images as test data with 12100 variables;  
Corresponding labels datasets:   
**Volcano?** : if in the image there are volcanoes (Main target), 1 (yes) or 0 (no)  
(If Volcano? = 0, the following three categories would be "nan")  
**Type** : 1= definitely a volcano,2 =probably, 3= possibly, 4= only a pit is visible   
**Radius** : is the radius of the volcano in the center of the image, in pixels   
**Number Volcanoes** : The number of volcanoes in the image  


## Data Import:
Data downloaded from Kaggle were csv files, there are four data files in total. The four data files were imported into R : "train_images", "train_labels", "test_images", and "test_labels". Dimensions of the four data files are as follows:  
train_images : 7000 observations and 12100 variables  
train_labels : 7000 observations and 4 variables  
test_images : 2734 observations and 12100 variables  
test_labels : 2734 observations and 4 variables  
Initial observations of the four data files are printed as following (due to large number of variables, for the "_images" files, only first 18 variables are printed):   
```{r, echo=FALSE}
library(data.table)
location = "/Users/hannahpu/OneDrive - University of Illinois - Urbana/STAT432/Project/"
train.x = fread(paste0(location, "train_images.csv"), header = FALSE)
test.x = fread(paste0(location, "test_images.csv"), header = FALSE)
train.y = fread(paste0(location, "train_labels.csv"), header = TRUE)
test.y = fread(paste0(location, "test_labels.csv"), header = TRUE)
#train.x = read.csv(paste0(location, "train_images.csv"), header = FALSE)
#test.x = read.csv(paste0(location, "test_images.csv"), header = FALSE)
#train.y = read.csv(paste0(location, "train_labels.csv"), header = TRUE)
#test.y = read.csv(paste0(location, "test_labels.csv"), header = TRUE)
```

**1. "train_images"**
```{r}
head(train.x[,1:18])
```

**2. "train_labels"**
```{r}
head(train.y)
```

**3. "test_images"**
```{r}
head(test.x[,1:18])
```

**4. "test_labels"**
```{r}
head(test.y)
```

\newpage
The image of first observation   
```{r, echo=FALSE, fig.height=5, fig.width=5}
train.x1 <- sapply(train.x[1:10,], as.numeric)
m <- matrix(train.x1[1,], nrow = 110, byrow = TRUE)
image(m, col = grey((0:255)/255))
```

## Dataset Background:
Finding Volcanoes On Venus.  
Kaggle. https://www.kaggle.com/amantheroot/finding-volcanoes-on-venus/data

## Methods:  
In our project, we are interested in detecting the volcanoes on Venus by analyzing and classifying the images.  
- Through our project, we are planning to apply k means clustering method to classify the images of the volcanoes into different categories.   
- We will also utilize the linear discriminant analysis (LDA) and the quadratic discriminant analysis (QDA) in our project.    
Our ultimate goal is to find the best method and build the best model that performs the best classification and has the minimum mean squared error to classify the images and match up with our label. 

## Challenges:
- We are dealing with large datasets (roughly 400 MB in total).
- We will have data visualization by converting the pixel observations into images.
- We will learn volcano knowledge to help us facilitate the process of classifying the volcanoes on venus.
- If we have more time, we want to further identity the number of volcanoes in each image rather than simply detecting if volcano exists in an image.




